Our Title

### 2015-05-03 draft.tex
Added in James' edits to the fuel cycle section.

ISIL poses a more direct threat to its neighbors than to the US homeland. While many foreigners with Western passports have pledged allegiance to ISIL, they often do so with no intent to return, as illustrated by the common burning of passports once individuals reach ISIL territory. Additionally, ISIL must continue to expand, taking territory from “illegitimate” governments, in order to maintain its own legitimacy as a Caliphate.

@@

\subsection{???????? Andrew hasn't proofed this section for content yet, just compiled it. From skimming though, it needs some love, as the voice is neither scientific nor professional     ??????????}

@@

The most important step in the process, from a proliferation standpoint, involves enriching the uranium in \ce{^{235}U}. Natural uranium is typically 0.72\% \ce{^{235}U} \cite{Benedict1981}.  If uranium is destined for a power plant, it is typically enriched to 3-5\% \ce{^{235}U}. If on the other hand it is going to be used in a weapon, the uranium must be enriched to greater than approximately 90\% \ce{^{235}U} \cite{Moody2014}.

The only true difference in the process of producing uranium for use in a nuclear reactor or in a weapon is the amount of enrichment carried out. Most enrichment is done using gas centrifuges, however this require significant infrastructure that is almost definitely beyond the capabilities of ISIL \cite{Benedict1981}.  Furthermore even if ISIL were able to gain control over an enrichment facility, the length of time it would take them to enrich the uranium for a weapon would likely be long enough to allow other countries to intervene militarily. For example, if ISIL were to gain control of Iran's Natanz facility, it would take them almost two months to enrich the uranium necessary for a weapon \cite{WisconsinProjectonNuclearArmsControl2015}. During that time, there would be obvious tell-tail signs of ISIL attempting to enrich uranium, such as power at that site, without needing to monitor nuclear signatures. Thus the most significant piece of stopping ISIL from enriching uranium for a weapon is simply identifying enrichment facilities, which are difficult to hide, with in its territory.

While ISIL could theoretically use laser separation to enrich uranium, which is significantly more difficult to detect, this is very unlikely. Currently, there is only one prototype process under investigation and it's in Australia \cite{Moody2014}.  Thus, it seems unlikely that ISIL would be able to use this method to enrich uranium, and thus it does not need to be monitored for.

@@

While uranium appears in nature, plutonium is only available through irradiation in a nuclear reactor \cite{Benedict1981}. It is formed when \ce{^{238}U} undergoes the process known as neutron capture \cite{Benedict1981}.  After a double beta decay, \ce{^{239}Pu}, which can be used in nuclear weapons, is produced \cite{Duderstadt1976}.  While heavy water reactors do have a higher percentage of \ce{^{238}U}, and thus form \ce{^{239}Pu} at a faster rate, it is possible to also use a light water reactor to produce plutonium \cite{Moody2014}.  While ISIL could theoretically gain control of a commercial nuclear reactor for plutonium production, this seems incredibly unlikely. The only commercial nuclear reactors currently near their territory are located in Iran \cite{WorldNuclearAssociation2015}.  Furthermore, while countries like Saudi Arabia, Egypt, Turkey, and the UAE are all moving towards nuclear power, it will be many years before any of those countries have commercial plants \cite{WorldNuclearAssociation2015}.  While there are not many commercial reactors that ISIL could easily use to produce plutonium, there are many research reactors scattered throughout the Middle East \cite{WorldNuclearAssociation2015a}.  However, ISIL would need to make very significant territory gains for any of the closest research reactors to be in danger \cite{BBC2015}.  Thus, it seems unlikely that ISIL would be able to gain control of a nuclear reactor for plutonium production.

It should be noted that Syria may have had a reactor for the production of plutonium in the past. However, that site was bombed and destroyed by an Israeli air strike in 2007 \cite{WorldNuclearAssociation2015a}. It is not impossible that ISIL could obtain the plans for that design. That being said, construction on a nuclear power plant typically takes at least 4 years \cite{NuclearEnergyInstitute2015}.  Furthermore, nuclear reactors are incredibly difficult to hide, as seen by Israel's 2007 air raid against Syria \cite{WorldNuclearAssociation2015a}.  Thus, it seems unlikely that ISIL would be able to complete such a reactor. Therefore, there is a very low probability that ISIL could covertly produce it's own plutonium, and thus little need to monitor for reprocessing signals from ISIL territory.

>>>>

ISIL poses a more direct threat to its neighbors than to the U.S. homeland. While many foreigners with Western passports have pledged allegiance to ISIL, they often do so with no intent to return, as illustrated by the common burning of passports once individuals reach ISIL territory. Additionally, ISIL must continue to expand, taking territory from “illegitimate” governments, in order to maintain its own legitimacy as a Caliphate.

@@

@@

The most important step in the process, from a proliferation standpoint, involves enriching the uranium in \ce{^{235}U}. Natural uranium is typically 0.72\% \ce{^{235}U} \cite{Benedict1981}.  If uranium is destined for a power plant, it is typically enriched to  3-5\% \ce{^{235}U}. For use in a nuclear weapon, the uranium must be enriched to greater than approximately 90\% \ce{^{235}U} \cite{Moody2014}.

Most enrichment is done using gas centrifuges; however, this requires significant infrastructure that is almost definitely beyond the capabilities of ISIL \cite{Benedict1981}.  Furthermore, even if ISIL were able to gain control over an enrichment facility, the length of time it would take them to enrich enough uranium for a weapon would allow other countries to intervene militarily. For example, if ISIL were to gain control of Iran's Natanz facility, it would take them approximately two (using 3.5\% enriched feedstock) to six months (using natural uranium feedstock) to enrich the uranium necessary for a weapon \cite{WisconsinProjectonNuclearArmsControl2015,Heinonen2015}. Therefore, the most significant step  in stopping ISIL from enriching uranium for a weapon is simply identifying enrichment facilities, which are difficult to hide, within or close to its territory.

While ISIL could theoretically use laser separation to enrich uranium, which is significantly more difficult to detect, this is very unlikely. Currently, there is only one prototype process under investigation, located in Australia, making it unlikely that ISIL would use this method to enrich uranium \cite{Moody2014}.

It is also possible that ISIL could attempt to use electromagnetic isotope separation (EMIS) to produce weapons usable material like the U.S. during World War II and Iraq before the First Gulf War. However, this seems incredulity unlikely. EMIS requires the use of machines called calutrons which are very expensive and very inefficient. For example, even with hundreds of calutrons functioning together, it still took the U.S. approximately one year to produce the uranium for the Hiroshima bomb \cite{Moody2014}. Thus, given the extreme inefficiency and lengthy enrichment time frame, it seems unlikely that ISIL would be able to construct and enrich uranium without the U.S. becoming aware and taking military action.

While gaseous diffusion is still used on a commercial scale, it does not pose much of a proliferation threat. It is an incredibly energy intensive process, requiring 50 times the power per separative unit required by the use of centrifuges \cite{WorldNuclearAssociation2015b}. If ISIL were to try and enrich uranium through gaseous diffusion, it would not be difficult for U.S. surveillance to find and destroy such a plant before they acquire enough highly enriched uranium.

One final way ISIL could attempt to enrich uranium is through chemical exchange. However, while this method has been developed in both France and Japan, it has never actually been used \cite{Africa2012}. Furthermore, this technique is not yet competitive with the other techniques discussed above, making it an unlikely choice \cite{Moody2014}.

@@

While uranium appears in nature, significant quantities of plutonium are only available through irradiation in a nuclear reactor \cite{Benedict1981}. It is formed when \ce{^{238}U} undergoes the process known as neutron capture \cite{Benedict1981}. \ce{^{239}Pu}, which can be used in nuclear weapons, is produced following a double beta decay of \ce{^{238}U} \cite{Duderstadt1976}. Heavy water reactors have a higher percentage of \ce{^{238}U}, thus forming \ce{^{239}Pu} at a faster rate, and are preferred to light water reactors to produce plutonium \cite{Moody2014}. While ISIL could theoretically gain control of a commercial nuclear reactor for plutonium production, it is deemed incredibly unlikely. The only commercial nuclear reactors currently near their territory are located in Iran \cite{WorldNuclearAssociation2015}. While countries like Saudi Arabia, Egypt, Turkey, and the UAE are all moving towards nuclear power, it will be many years before any of those countries have commercial plants \cite{WorldNuclearAssociation2015}. While there are not many commercial reactors that ISIL could easily use to produce plutonium, there are many research reactors scattered throughout the Middle East \cite{WorldNuclearAssociation2015a}. However, ISIL would need to make very significant territory gains for any of the closest research reactors to be in danger \cite{BBC2015}. Thus, it seems unlikely that ISIL would be able to gain control of a nuclear reactor for plutonium production. Additionally, any captured reactor fuel would require reprocessing to separate out the plutonium. The time required would be significant enough to allow for a response.

It should be noted that Syria may have had a reactor for the production of plutonium in the past. However, that site was bombed and destroyed by an Israeli airstrike in 2007 \cite{WorldNuclearAssociation2015a}. It is not impossible that ISIL could obtain the plans for that design. Nevertheless, construction on a nuclear power plant typically takes at least 4 years, and nuclear reactors are incredibly difficult to hide \cite{WorldNuclearAssociation2015a,NuclearEnergyInstitute2015}. Therefore, there is a very low probability that ISIL could covertly produce its own plutonium, making this route of proliferation unlikely.
